Assessment of the KLGO Coastal Waterbird Survey Monitoring Program

### 2019-02-13 analysis/paper/paper.Rmd
Fix some biblio errors; initial tests of figure labels and captions.

---
title: "Assessment of the KLGO Coastal Waterbird Survey Monitoring Program"
author:
- name: Joel H. Reynolds^[U.S. National Park Service, Alaska Region, Natural Resources
    Science Team]
- name: Madeleine Ward^[Swarthmore College, Dept. of Mathematics and Statistics]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
keywords: keyword 1; keyword 2; keyword 3
link-citations: yes
bibliography: KLGOCWS_References.bib
abstract: The Coastal Waterbird Survey has been conducted, with varying levels of
  effort and in-season frequency, since 2003 [@Surdyk2018]. The area of interest has
  been divided into eight contiguous survey units; on each survey, at each unit the
  number, age, composition (sex), breeding status, and noteworthy behavioral observations
  are recorded for each species [@Anonymous2009]. Given the constraints on personnel,
  time, and budgets, it is imperative that surveys be conducted efficiently in terms
  of both choice of sampling design and choice of sampling effort. Now that KLGO has
  multiple years of experience implementing the protocol, it is timely to step back
  and assess how things are going. We review the current program's objectives and survey design,
  identify statistical issues, and recommend potential solutions. Major topics include... blah blah. The study concludes with a series of recommended tasks that the park should
  undertake to improve the coastal waterbird survey's efficiency and effectiveness.
---

@@

# Issues to Resolve
- How do we use latex figure environment in rmarkdown (for inclusion of ggplot graphics; captioning figures and labelling for cross-references, etc)?
- conversion from paper.md to paper.pdf (pandoc issue?) 
  (i) update MikTex package, set dev=c('pdf') in setup
- github permissions for Madeleine on local machine
- adding files to .gitignore
- What is up with 2008 surveys? why such distinct timing?

# Introduction
Since 2003, staff at Klondike Gold Rush National Historical Park (KLGO) have surveyed coastal waterbird species distribution and abundance multiple times each year from late April through late September (Coastal Waterbird Survey or ‘CWS’; [@Surdyk2018]).  This is sufficient history to warrant stepping back and assessing the monitoring program.  Such assessments are key to an adaptive monitoring program [@Lindenmeyer 2009], are an important and distinct phase to promote program, and institutional, learning and evolution [@Reynolds2016], and align with guidance from NPS' Inventory and Monitoring Division [@Gallo2018a, @Mitchell2018].

@@

```{r Data_Retrieval, eval=TRUE, echo=FALSE, warning = FALSE}
# if (system("git config user.email")=="mward2@swarthmore.edu") 
#      source("./scripts/data_retrievalSwarthmore.R")
# else source("./scripts/data_retrievalNPS.R")

@@

```{r EDA_SurveyTiming, eval=TRUE, cache = FALSE, fig.show= 'asis'}
 source("../scripts/survey_timing_EDA.R", echo=FALSE) 
print(Plot1)

@@

>>>>

---
title: "Assessment of the KLGO Coastal Waterbird Survey Monitoring Program"
author:
- name: Joel H. Reynolds^[U.S. National Park Service, Alaska Region, Natural Resources
    Science Team]
- name: Madeleine Ward^[Swarthmore College, Dept. of Mathematics and Statistics]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::word_document2:
    fig_caption: yes
    toc: yes
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    fig_caption: yes
    toc: yes
keywords: keyword 1; keyword 2; keyword 3
link-citations: yes
bibliography: KLGOCWS_References.bib
abstract: The Coastal Waterbird Survey has been conducted, with varying levels of
  effort and in-season frequency, since 2003 [@Surdyk2018]. The area of interest has
  been divided into eight contiguous survey units; on each survey, at each unit the
  number, age, composition (sex), breeding status, and noteworthy behavioral observations
  are recorded for each species [@Anonymous2009]. Given the constraints on personnel,
  time, and budgets, it is imperative that surveys be conducted efficiently in terms
  of both choice of sampling design and choice of sampling effort. Now that KLGO has
  multiple years of experience implementing the protocol, it is timely to step back
  and assess how things are going. We review the current program's objectives and
  survey design, identify statistical issues, and recommend potential solutions. Major
  topics include... blah blah. The study concludes with a series of recommended tasks
  that the park should undertake to improve the coastal waterbird survey's efficiency
  and effectiveness.
---

@@

# Issues to Resolve with Rmarkdown, Knitr, Github, etc.
- conversion from paper.md to paper.pdf (pandoc issue?) 
  (i) update MikTex package, set dev=c('pdf') in setup
- github permissions for Madeleine on local machine
- adding files to .gitignore

# Issues to Resolve with content (narrative, figures, analyses)
- What is up with 2008 surveys? why such distinct timing?

# Introduction
Since 2003, staff at Klondike Gold Rush National Historical Park (KLGO) have surveyed coastal waterbird species distribution and abundance multiple times each year from late April through late September (Coastal Waterbird Survey or ‘CWS’; [@Surdyk2018]).  This is sufficient history to warrant stepping back and assessing the monitoring program.  Such assessments are key to an adaptive monitoring program [@Lindenmayer2009], are an important and distinct phase to promote program, and institutional, learning and evolution [@Reynolds2016], and align with guidance from NPS' Inventory and Monitoring Division [@Gallo2018a, @Mitchell2018].

@@

```{r Data-Retrieval, eval=TRUE, echo=FALSE, warning = FALSE}
# if (system("git config user.email")=="mward2@swarthmore.edu") 
#      source("./scripts/data_retrievalSwarthmore.R")
# else source("./scripts/data_retrievalNPS.R")

@@

The annual dates of the first and last survey each year has varied by over a month (Figure \@ref(fig:Fig-SurveyTiming)).

```{r EDA_SurveyTiming, eval=TRUE, cache = TRUE, include = FALSE}
# dotplot of survey times by year
source("../scripts/survey_timing_EDA.R", echo=FALSE)

@@

```{r Fig-SurveyTiming, echo=FALSE, eval=TRUE, cache=TRUE, dependson= c("EDA_SurveyTiming"), fig.show= 'asis', fig.cap="The annual dates of the first and last survey have both varied by over a month across the years."}

print(Plot1)
```
